---
title: 'Markov Sick-Sicker model in R'
subtitle: "With simulation-time dependency"
author: "Wang Ye"
output:
  pdf_document: default
  html_document: default
keep_tex: yes
self_contained: no
---
Authors: 

- Wang Ye  <1589936809@qq.com>

- Wang Hao

In collaboration of: 		

1. Nanjing Drum Tower Hospital, China Pharmaceutical University, Nanjing, China

2. Department of Pharmacy, Drum Tower Hospital Affiliated to Medical School of 
Nanjing University, Nanjing, China

This Code adaptation Developed by the Decision Analysis in R for Technologies in Health (DARTH) workgroup.

This code implements a simulation-time-dependent Sick-Sicker cSTM model to  conduct a CEA of two strategies:
- Standard of Care (SoC): best available care for the patients with the disease. This scenario reflects the natural history of the disease  
  progression.
- Strategy AB: This strategy combines treatment A and treatment B. The disease  progression is reduced, and individuals in the Sick state have an  
  improved quality of life.

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval = TRUE)
```

Change `eval` to `TRUE` if you want to knit this document.

```{r}
rm(list = ls())      # clear memory (removes all the variables from the workspace)
```

# 01 Load packages

```{r, warning = F, message = F}
if (!require('pacman')) install.packages('pacman'); library(pacman) # use this package to conveniently install other packages
# load (install if required) packages from CRAN
p_load("dplyr", "tidyr", "reshape2", "devtools", "scales", "ellipse", "ggplot2", "ggrepel", "gridExtra", "lazyeval", "igraph", "truncnorm", "ggraph", "reshape2", "patchwork", "knitr", "stringr", "diagram", "dampack")                                               
# load (install if required) packages from GitHub
# install_github("DARTH-git/darthtools", force = TRUE) #Uncomment if there is a newer version
p_load_gh("DARTH-git/darthtools")
```

# 02 Load functions

```{r}
# all functions are in the darthtools package
```

# 03 Model input

```{r}
## General setup 
cycle_length <- 1   # cycle length equal to one year (use 1/12 for monthly)
n_age_init   <- 25  # age at baseline
n_age_max    <- 100 # maximum age of follow up
n_cycles     <- (n_age_max - n_age_init)/cycle_length # time horizon, number of cycles
# Age labels 
v_age_names  <- paste(rep(n_age_init:(n_age_max-1), each = 1/cycle_length), 
                      1:(1/cycle_length), 
                      sep = ".")
# the 4 health states of the model:
v_names_states <- c("H",  # Healthy (H)
                    "S1", # Sick (S1)
                    "S2", # Sicker (S2)
                    "D")  # Dead (D)
                                           
n_states <- length(v_names_states)   # number of health states 

### Discounting factors 
d_c <- 0.03 # annual discount rate for costs 
d_e <- 0.03 # annual discount rate for QALYs

### Strategies 
v_names_str <- c("Standard of care", # store the strategy names
                 "Strategy AB") 
n_str       <- length(v_names_str)   # number of strategies

## Within-cycle correction (WCC) using half-cycle rule 
v_wcc  <- gen_wcc(n_cycles = n_cycles, method = "half-cycle")

### Transition rates (annual), and hazard ratios (HRs) 
r_HS1  <- 0.15  # constant annual rate of becoming Sick when Healthy
r_S1H  <- 0.5   # constant annual rate of becoming Healthy when Sick
r_S1S2 <- 0.105 # constant annual rate of becoming Sicker when Sick
hr_S1  <- 3     # hazard ratio of death in Sick vs Healthy 
hr_S2  <- 10    # hazard ratio of death in Sicker vs Healthy 

### Effectiveness of treatment AB 
hr_S1S2_trtAB <- 0.6  # hazard ratio of becoming Sicker when Sick under treatment AB

## Age-dependent mortality rates 
lt_usa_2015 <- read.csv("HMD_USA_Mx_2015.csv")
# Extract age-specific all-cause mortality for ages in model time horizon
v_r_mort_by_age <- lt_usa_2015 %>% 
  dplyr::filter(Age >= n_age_init & Age < n_age_max) %>%
  dplyr::select(Total) %>%
  as.matrix()

### State rewards 
#### Costs 
c_H     <- 2000  # annual cost of being Healthy
c_S1    <- 4000  # annual cost of being Sick
c_S2    <- 15000 # annual cost of being Sicker
c_D     <- 0     # annual cost of being dead
c_trtAB <- 25000 # annual cost of receiving treatment AB
#### Utilities 
u_H     <- 1     # annual utility of being Healthy
u_S1    <- 0.75  # annual utility of being Sick
u_S2    <- 0.5   # annual utility of being Sicker
u_D     <- 0     # annual utility of being dead
u_trtAB <- 0.95  # annual utility when receiving treatment AB

### Transition rewards 
du_HS1  <- 0  # disutility when transitioning from Healthy to Sick
ic_HS1  <- 0  # increase in cost when transitioning from Healthy to Sick
ic_D    <- 0  # increase in cost when dying

### Discount weight for costs and effects 
v_dwc   <- 1 / ((1 + (d_e * cycle_length)) ^ (0:n_cycles))
v_dwe   <- 1 / ((1 + (d_c * cycle_length)) ^ (0:n_cycles))

# Process model inputs 
## Age-specific transition rates to the Dead state for all cycles 
v_r_HDage  <- rep(v_r_mort_by_age, each = 1/cycle_length)
# Name age-specific mortality vector 
names(v_r_HDage) <- v_age_names

# compute mortality rates
v_r_S1Dage  <- v_r_HDage * hr_S1 # Age-specific mortality rate in the Sick state 
v_r_S2Dage  <- v_r_HDage * hr_S2 # Age-specific mortality rate in the Sicker state 
v_r_S2Dage

# transform rates to probabilities adjusting by cycle length
p_HS1       <- rate_to_prob(r = r_HS1,  t = cycle_length) # constant annual probability of becoming Sick when Healthy conditional on surviving 
p_S1H       <- rate_to_prob(r = r_S1H,  t = cycle_length) # constant annual probability of becoming Healthy when Sick conditional on surviving
p_S1S2      <- rate_to_prob(r = r_S1S2, t = cycle_length) # constant annual probability of becoming Sicker when Sick conditional on surviving
v_p_HDage   <- rate_to_prob(v_r_HDage,  t = cycle_length) # Age-specific mortality risk in the Healthy state 
v_p_S1Dage  <- rate_to_prob(v_r_S1Dage, t = cycle_length) # Age-specific mortality risk in the Sick state
v_p_S2Dage  <- rate_to_prob(v_r_S2Dage, t = cycle_length) # Age-specific mortality risk in the Sicker state
v_p_S2Dage
## Annual transition probability of becoming Sicker when Sick for treatment AB 
# Apply hazard ratio to rate to obtain transition rate of becoming Sicker when Sick for treatment AB
r_S1S2_trtAB <- r_S1S2 * hr_S1S2_trtAB
# Transform rate to probability to become Sicker when Sick under treatment AB 
# adjusting by cycle length conditional on surviving
p_S1S2_trtAB <- rate_to_prob(r = r_S1S2_trtAB, t = cycle_length)
```

# 04 Construct state-transition models

## 04.1 Initial state vector

```{r}
# All starting healthy
v_m_init <- c(H = 1, S1 = 0, S2 = 0, D = 0) # initial state vector
v_m_init
```

## 04.2 Initialize cohort traces

```{r}
### Initialize cohort trace under SoC 
m_M_SoC <- matrix(NA, 
              nrow = (n_cycles + 1), ncol = n_states, 
              dimnames = list(0:n_cycles, v_names_states))
# Store the initial state vector in the first row of the cohort trace
m_M_SoC[1, ] <- v_m_init

### Initialize cohort trace for strategy AB 
# Structure and initial states are the same as for SoC
m_M_strAB <- m_M_SoC # Strategy AB
```

## 04.3 Create transition probability matrices

```{r}
## Create transition probability arrays for strategy SoC 
### Initialize transition probability array for strategy SoC 
# All transitions to a non-death state are assumed to be conditional on survival
a_P_SoC <- array(0,
                 dim  = c(n_states, n_states, n_cycles),
                 dimnames = list(v_names_states, 
                                 v_names_states, 
                                 0:(n_cycles - 1)))
### Fill in array
## From H
a_P_SoC["H", "H", ]   <- (1 - v_p_HDage) * (1 - p_HS1)
a_P_SoC["H", "S1", ]  <- (1 - v_p_HDage) *      p_HS1
a_P_SoC["H", "D", ]   <-      v_p_HDage
## From S1
a_P_SoC["S1", "H", ]  <- (1 - v_p_S1Dage) *       p_S1H
a_P_SoC["S1", "S1", ] <- (1 - v_p_S1Dage) * (1 - (p_S1H + p_S1S2))
a_P_SoC["S1", "S2", ] <- (1 - v_p_S1Dage) *               p_S1S2
a_P_SoC["S1", "D", ]  <-      v_p_S1Dage
## From S2
a_P_SoC["S2", "S2", ] <- 1 - v_p_S2Dage
a_P_SoC["S2", "D", ]  <-     v_p_S2Dage
## From D
a_P_SoC["D", "D", ]   <- 1

### Initialize transition probability array for strategy AB 
a_P_strAB <- a_P_SoC
# Update only transition probabilities from S1 involving p_S1S2
a_P_strAB["S1", "S1", ] <- (1 - v_p_S1Dage) * (1 - (p_S1H + p_S1S2_trtAB))
a_P_strAB["S1", "S2", ] <- (1 - v_p_S1Dage) *               p_S1S2_trtAB

## Check if transition probability arrays are valid 
### Check that transition probabilities are [0, 1] 
check_transition_probability(a_P_SoC,   verbose = TRUE)
check_transition_probability(a_P_strAB, verbose = TRUE)
### Check that all rows for each slice of the array sum to 1 
check_sum_of_transition_array(a_P_SoC,   n_states = n_states, n_cycles = n_cycles, verbose = TRUE)
check_sum_of_transition_array(a_P_strAB, n_states = n_states, n_cycles = n_cycles, verbose = TRUE)
```

## 04.4 Create transition dynamics arrays

```{r}
# These arrays will capture transitions from each state to another over time 
### Initialize transition dynamics array for strategy SoC 
a_A_SoC <- array(0,
             dim      = c(n_states, n_states, n_cycles + 1),
             dimnames = list(v_names_states, v_names_states, 0:n_cycles))
# Set first slice of a_A_SoC with the initial state vector in its diagonal
diag(a_A_SoC[, , 1]) <- v_m_init
### Initialize transition-dynamics array for strategy AB 
# Structure and initial states are the same as for SoC
a_A_strAB <- a_A_SoC
```

# 05 Run Markov model

```{r}
# Iterative solution of age-dependent cSTM
for(t in 1:n_cycles){
  ## Fill in cohort trace
  # For SoC
  m_M_SoC[t + 1, ]   <- m_M_SoC[t, ]   %*% a_P_SoC[, , t]
  # For strategy AB 
  m_M_strAB[t + 1, ] <- m_M_strAB[t, ] %*% a_P_strAB[, , t]
  
  ## Fill in transition-dynamics array
  # For SoC
  a_A_SoC[, , t + 1]   <- diag(m_M_SoC[t, ])   %*% a_P_SoC[, , t]
  # For strategy AB
  a_A_strAB[, , t + 1] <- diag(m_M_strAB[t, ]) %*% a_P_strAB[, , t]
}

## Store the cohort traces in a list 
l_m_M <- list(SoC =  m_M_SoC,
              AB  =  m_M_strAB)
names(l_m_M) <- v_names_str

## Store the transition dynamics array for each strategy in a list 
l_a_A <- list(SoC =  a_A_SoC,
              AB  =  a_A_strAB)
names(l_a_A) <- v_names_str
```

# 06 Plot Outputs

## 06.1 Plot the cohort trace for strategies SoC and AB

```{r}
plot_trace(m_M_SoC)
plot_trace(m_M_strAB)
```

# 07 State Rewards 

```{r}
## Scale by the cycle length 
# Vector of state utilities under strategy SoC
v_u_SoC    <- c(H  = u_H, 
                S1 = u_S1, 
                S2 = u_S2, 
                D  = u_D) * cycle_length
# Vector of state costs under strategy SoC
v_c_SoC    <- c(H  = c_H, 
                S1 = c_S1,
                S2 = c_S2, 
                D  = c_D) * cycle_length
# Vector of state utilities under strategy AB
v_u_strAB  <- c(H  = u_H, 
                S1 = u_trtAB, 
                S2 = u_S2, 
                D  = u_D) * cycle_length
# Vector of state costs under strategy AB
v_c_strAB  <- c(H  = c_H, 
                S1 = c_S1 + c_trtAB, 
                S2 = c_S2 + c_trtAB, 
                D  = c_D) * cycle_length

## Store state rewards 
# Store the vectors of state utilities for each strategy in a list 
l_u <- list(SoC = v_u_SoC,
            AB  = v_u_strAB)
# Store the vectors of state cost for each strategy in a list 
l_c <- list(SoC =  v_c_SoC,
            AB  =  v_c_strAB)

# assign strategy names to matching items in the lists
names(l_u) <- names(l_c) <- v_names_str
```

# 08 Compute expected outcomes 

```{r}
# Create empty vectors to store total utilities and costs 
v_tot_qaly <- v_tot_cost <- vector(mode = "numeric", length = n_str)
names(v_tot_qaly) <- names(v_tot_cost) <- v_names_str

## Loop through each strategy and calculate total utilities and costs 
for (i in 1:n_str) { # i <- 1
  v_u_str <- l_u[[i]]   # select the vector of state utilities for the i-th strategy
  v_c_str <- l_c[[i]]   # select the vector of state costs for the i-th strategy
  a_A_str <- l_a_A[[i]] # select the transition array for the i-th strategy
  
  ## Array of state rewards 
  # Create transition matrices of state utilities and state costs for the i-th strategy 
  m_u_str   <- matrix(v_u_str, nrow = n_states, ncol = n_states, byrow = T)
  m_c_str   <- matrix(v_c_str, nrow = n_states, ncol = n_states, byrow = T)
  # Expand the transition matrix of state utilities across cycles to form a transition array of state utilities
  a_R_u_str <- array(m_u_str, 
                     dim      = c(n_states, n_states, n_cycles + 1),
                     dimnames = list(v_names_states, v_names_states, 0:n_cycles))
  # Expand the transition matrix of state costs across cycles to form a transition array of state costs
  a_R_c_str <- array(m_c_str, 
                     dim      = c(n_states, n_states, n_cycles + 1),
                     dimnames = list(v_names_states, v_names_states, 0:n_cycles))
  
  ## Apply transition rewards
  # Apply disutility due to transition from H to S1
  a_R_u_str["H", "S1", ]      <- a_R_u_str["H", "S1", ]       - du_HS1
  # Add transition cost per cycle due to transition from H to S1
  a_R_c_str["H", "S1", ]      <- a_R_c_str["H", "S1", ]       + ic_HS1
  # Add transition cost  per cycle of dying from all non-dead states
  a_R_c_str[-n_states, "D", ] <- a_R_c_str[-n_states, "D", ] + ic_D
  
  ### Expected QALYs and costs for all transitions per cycle
  # QALYs = life years x QoL
  # Note: all parameters are annual in our example. In case your own case example is different make sure you correctly apply.
  a_Y_c_str <- a_A_str * a_R_c_str
  a_Y_u_str <- a_A_str * a_R_u_str 
  
  ### Expected QALYs and costs per cycle
  ## Vector of QALYs and costs
  v_qaly_str <- apply(a_Y_u_str, 3, sum) # sum the proportion of the cohort across transitions 
  v_cost_str <- apply(a_Y_c_str, 3, sum) # sum the proportion of the cohort across transitions
  
  ## Discounted total expected QALYs and Costs per strategy and apply within-cycle correction if applicable
  # QALYs
  v_tot_qaly[i] <- t(v_qaly_str) %*% (v_dwe * v_wcc)
  # Costs
  v_tot_cost[i] <- t(v_cost_str) %*% (v_dwc * v_wcc)
}
```

# 09 Cost-effectiveness analysis (CEA) 

```{r}
## Incremental cost-effectiveness ratios (ICERs) 
df_cea <- calculate_icers(cost       = v_tot_cost, 
                          effect     = v_tot_qaly,
                          strategies = v_names_str)
df_cea
```

```{r}
## CEA table in proper format 
table_cea <- format_table_cea(df_cea)
table_cea
```

```{r}
## CEA frontier 
plot(df_cea, label = "all", txtsize = 16) +
  expand_limits(x = max(table_cea$QALYs) + 0.1) +
  theme(legend.position = c(0.82, 0.3))
```

# 10 Deterministic Sensitivity Analysis (DSA)

```{r}
## Load model, CEA and PSA functions 
source('Functions_markov_sick-sicker_time.R')
```

## 10.1 Model input for SA

```{r}
## List of input parameters 
l_params_all <- list(
  # Transition probabilities (per cycle), hazard ratios
  v_r_HDage = v_r_HDage, # constant rate of dying when Healthy (all-cause mortality)
  r_HS1     = 0.15,      # constant annual rate of becoming Sick when Healthy conditional on surviving
  r_S1H     = 0.5,       # constant annual rate of becoming Healthy when Sick conditional on surviving
  r_S1S2    = 0.105,     # constant annual rate of becoming Sicker when Sick conditional on surviving
  hr_S1     = 3,         # hazard ratio of death in Sick vs Healthy 
  hr_S2     = 10,        # hazard ratio of death in Sicker vs Healthy 
  # Effectiveness of treatment AB 
  hr_S1S2_trtAB = 0.6,   # hazard ratio of becoming Sicker when Sick under treatment AB
  ## State rewards
  # Costs
  c_H       = 2000,      # cost of remaining one cycle in Healthy 
  c_S1      = 4000,      # cost of remaining one cycle in Sick 
  c_S2      = 15000,     # cost of remaining one cycle in Sicker 
  c_D       = 0,         # cost of being dead (per cycle)
  c_trtAB   = 25000,     # cost of treatment A
  # Utilities
  u_H       = 1,         # utility when Healthy 
  u_S1      = 0.75,      # utility when Sick 
  u_S2      = 0.5,       # utility when Sicker
  u_D       = 0,         # utility when Dead 
  u_trtAB   = 0.95,      # utility when being treated with A
  ## Transition rewards
  du_HS1    = 0.01,      # disutility when transitioning from Healthy to Sick
  ic_HS1    = 1000,      # increase in cost when transitioning from Healthy to Sick
  ic_D      = 2000,      # increase in cost when dying
  # Initial and maximum ages
  n_age_init = 25,
  n_age_max  = 100,
  # Discount rates
  d_c        = 0.03,     # annual discount rate for costs 
  d_e        = 0.03,     # annual discount rate for QALYs,
  # Cycle length
  cycle_length = 1
)
```

## 10.2 One-way sensitivity analysis (OWSA)

```{r}
options(scipen = 999) # disabling scientific notation in R
# data.frame containing all parameters, their base-case values, and the min and 
# max values of the parameters of interest 
df_params_owsa <- data.frame(pars = c("r_S1S2", "c_trtAB", "u_S1", "u_trtAB"),
                             min  = c(0.05 , 18000 , 0.65, 0.80), # min parameter values
                             max  = c(0.155, 36000 , 0.85, 0.98)  # max parameter values
                             )

owsa_nmb  <- run_owsa_det(params_range    = df_params_owsa,   # data.frame with parameters for OWSA
                          params_basecase = l_params_all,     # list with all parameters
                          nsamp           = 100,              # number of parameter values
                          FUN             = calculate_ce_out, # function to compute outputs
                          outcomes        = c("NMB"),         # output to do the OWSA on
                          strategies      = v_names_str,      # names of the strategies
                          n_wtp           = 120000)           # extra argument to pass to FUN
```

```{r}
plot(owsa_nmb, txtsize = 10, n_x_ticks = 4, 
     facet_scales = "free") +
     theme(legend.position = "bottom")
```

### 10.2.1 Optimal strategy with OWSA

```{r}
owsa_opt_strat(owsa = owsa_nmb, txtsize = 10)
```

### 10.2.2 Tornado plot

```{r}
owsa_tornado(owsa = owsa_nmb)
```

## 10.3 Two-way sensitivity analysis (TWSA)

```{r}
# dataframe containing all parameters, their basecase values, and the min and 
# max values of the parameters of interest
df_params_twsa <- data.frame(pars = c("c_trtAB", "u_trtAB"),
                             min  = c(18000, 0.80),  # min parameter values
                             max  = c(36000, 0.98)  # max parameter values
                             )

twsa_nmb <- run_twsa_det(params_range    = df_params_twsa,   # data.frame with parameters for TWSA
                         params_basecase = l_params_all,     # list with all parameters
                         nsamp           = 40,               # number of parameter values
                         FUN             = calculate_ce_out, # function to compute outputs
                         outcomes        = c("NMB"),         # output to do the TWSA on
                         strategies      = v_names_str,      # names of the strategies
                         n_wtp           = 120000)           # extra argument to pass to FUN
```

### 10.3.1 Plot TWSA

```{r}
plot(twsa_nmb)
```

# 11 Probabilistic Sensitivity Analysis (PSA) 

## 11.1 Model input

```{r}
# Store the parameter names into a vector
v_names_params <- names(l_params_all)

## Test functions to generate CE outcomes and PSA dataset 
# Test function to compute CE outcomes
calculate_ce_out(l_params_all) 

# Test function to generate PSA input dataset
generate_psa_params(10) 

## Generate PSA dataset 
# Number of simulations
n_sim <- 1000

# Generate PSA input dataset
df_psa_input <- generate_psa_params(n_sim = n_sim)
# First six observations
head(df_psa_input)

### Histogram of parameters 
ggplot(melt(df_psa_input, variable.name = "Parameter"), aes(x = value)) +
  facet_wrap(~Parameter, scales = "free") +
  geom_histogram(aes(y = ..density..)) +
  ylab("") +
  theme_bw(base_size = 16) + 
  theme(axis.text = element_text(size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank()) 
```

## 11.2 Run PSA

```{r}
# Initialize data.frames with PSA output 
# data.frame of costs
df_c <- as.data.frame(matrix(0, 
                             nrow = n_sim,
                             ncol = n_str))
colnames(df_c) <- v_names_str
# data.frame of effectiveness
df_e <- as.data.frame(matrix(0, 
                             nrow = n_sim,
                             ncol = n_str))
colnames(df_e) <- v_names_str

# Conduct probabilistic sensitivity analysis
# Run Markov model on each parameter set of PSA input dataset
n_time_init_psa_series <- Sys.time()
for (i in 1:n_sim) { # i <- 1
  l_psa_input <- update_param_list(l_params_all, df_psa_input[i,])
  # Outcomes
  l_out_ce_temp  <- calculate_ce_out(l_psa_input)
  df_c[i, ]  <- l_out_ce_temp$Cost  
  df_e[i, ]  <- l_out_ce_temp$Effect
  # Display simulation progress
  if (i/(n_sim/100) == round(i/(n_sim/100), 0)) { # display progress every 5%
    cat('\r', paste(i/n_sim * 100, "% done", sep = " "))
  }
}
n_time_end_psa_series <- Sys.time()
n_time_total_psa_series <- n_time_end_psa_series - n_time_init_psa_series
print(paste0("PSA with ", scales::comma(n_sim), " simulations run in series in ", 
             round(n_time_total_psa_series, 2), " ", 
             units(n_time_total_psa_series)))
```

# 11.3 Visualize PSA results for CEA 

```{r}
### Create PSA object 
l_psa <- make_psa_obj(cost          = df_c, 
                      effectiveness = df_e, 
                      parameters    = df_psa_input, 
                      strategies    = v_names_str)
l_psa$strategies <- v_names_str
colnames(l_psa$effectiveness) <- v_names_str
colnames(l_psa$cost) <- v_names_str

# Vector with willingness-to-pay (WTP) thresholds.
v_wtp <- seq(0, 200000, by = 5000)
```

## 11.3.1 Cost-Effectiveness Scatter plot 

```{r}
### Cost-Effectiveness Scatter plot 
txtsize <- 13
gg_scattter <- plot_psa(l_psa, txtsize = txtsize) +
  ggthemes::scale_color_colorblind() +
  ggthemes::scale_fill_colorblind() +
  scale_y_continuous("Cost (Thousand $)", 
                     breaks = number_ticks(10),
                     labels = function(x) x/1000) +
  xlab("Effectiveness (QALYs)") +
  guides(col = guide_legend(nrow = 2)) +
  theme(legend.position = "bottom")
gg_scattter
```

## 11.3.2 Incremental cost-effectiveness ratios (ICERs) with probabilistic output

```{r}
### Incremental cost-effectiveness ratios (ICERs) with probabilistic output 
# Compute expected costs and effects for each strategy from the PSA
df_out_ce_psa <- summary(l_psa)
df_cea_psa <- calculate_icers(cost       = df_out_ce_psa$meanCost, 
                              effect     = df_out_ce_psa$meanEffect,
                              strategies = df_out_ce_psa$Strategy)
df_cea_psa
```

## 11.3.3 Plot cost-effectiveness frontier with probabilistic output

```{r}
### Plot cost-effectiveness frontier with probabilistic output 
plot_icers(df_cea_psa, label = "all", txtsize = txtsize) +
  expand_limits(x = max(table_cea$QALYs) + 0.1) +
  theme(legend.position = c(0.8, 0.3))
```

## 11.3.4 Cost-effectiveness acceptability curves (CEACs) and frontier (CEAF)

```{r}
### Cost-effectiveness acceptability curves (CEACs) and frontier (CEAF) 
ceac_obj <- ceac(wtp = v_wtp, psa = l_psa)
# Regions of highest probability of cost-effectiveness for each strategy
summary(ceac_obj)
# CEAC & CEAF plot
gg_ceac <- plot_ceac(ceac_obj, txtsize = txtsize, xlim = c(0, NA), n_x_ticks = 14) +
  ggthemes::scale_color_colorblind() +
  ggthemes::scale_fill_colorblind() +
  theme(legend.position = c(0.8, 0.48))
gg_ceac
```

## 11.3.5 Expected Loss Curves (ELCs)

```{r}
### Expected Loss Curves (ELCs) 
elc_obj <- calc_exp_loss(wtp = v_wtp, psa = l_psa)
elc_obj

# ELC plot
gg_elc <- plot_exp_loss(elc_obj, log_y = FALSE, 
               txtsize = txtsize, xlim = c(0, NA), n_x_ticks = 14,
               col = "full") +
  ggthemes::scale_color_colorblind() +
  ggthemes::scale_fill_colorblind() +
  # geom_point(aes(shape = as.name("Strategy"))) +
  scale_y_continuous("Expected Loss (Thousand $)", 
                     breaks = number_ticks(10),
                     labels = function(x) x/1000) +
  theme(legend.position = c(0.4, 0.7),)
gg_elc
```

## 11.3.6 Expected value of perfect information (EVPI) 

```{r}
### Expected value of perfect information (EVPI) 
evpi <- calc_evpi(wtp = v_wtp, psa = l_psa)
# EVPI plot
gg_evpi <- plot_evpi(evpi, effect_units = "QALY", 
                     txtsize = txtsize, xlim = c(0, NA), n_x_ticks = 14) +
  scale_y_continuous("EVPI (Thousand $)", 
                     breaks = number_ticks(10),
                     labels = function(x) x/1000)
gg_evpi
```

# REFERENCES

- Alarid-Escudero F, Krijkamp EM, Enns EA, Yang A, Hunink MGM Pechlivanoglou P,
Jalal H. A Tutorial on Time-Dependent Cohort State-Transition Models in R using 
a Cost-Effectiveness Analysis Example. Medical Decision Making, 2022 (In press):
1-21. https://doi.org/10.1177/0272989X221121747

- Alarid-Escudero F, Krijkamp EM, Enns EA, Yang A, Hunink MGM Pechlivanoglou P,
Jalal H. An Introductory Tutorial on Cohort State-Transition Models in R Using a 
Cost-Effectiveness Analysis Example. Medical Decision Making, 2022 
(Online First):1-18. https://doi.org/10.1177/0272989X221103163

- Alarid-Escudero F, Krijkamp EM, Enns EA, Yang A, Hunink MGM Pechlivanoglou P,
Jalal H. An Introductory Tutorial on Cohort State-Transition Models in R Using a 
Cost-Effectiveness Analysis Example. Medical Decision Making, 2022 
(Online First):1-18. https://doi.org/10.1177/0272989X221103163

- Jalal H, Pechlivanoglou P, Krijkamp E, Alarid-Escudero F, Enns E, Hunink MG. 
An Overview of R in Health Decision Sciences. Med Decis Making. 2017; 37(3): 735-746. 
https://journals.sagepub.com/doi/abs/10.1177/0272989X16686559

- Krijkamp EM, Alarid-Escudero F, Enns EA, Jalal HJ, Hunink MGM, Pechlivanoglou P. 
Microsimulation modeling for health decision sciences using R: A tutorial. 
Med Decis Making. 2018;38(3):400–22. 
https://journals.sagepub.com/doi/abs/10.1177/0272989X18754513
 
- Krijkamp EM, Alarid-Escudero F, Enns E, Pechlivanoglou P, Hunink MM, Jalal H. 
A Multidimensional Array Representation of State-Transition Model Dynamics. 
Med Decis Mak. 2020;40(2):242-248. https://doi.org/10.1177/0272989X19893973
